---
title: "SharksFishCoralDAG"
author: "Simon Dedman"
format: html
editor: visual
---

```{r install library dagitty}
#| echo: false
# install.packages("dagitty")
remotes::install_github("jtextor/dagitty/r")
library(dagitty)
# https://dagitty.net/dags.html?id=JHVro8aM 2024-05-24
myDag <- dagitty('dag {
ambush_piscivore [pos="-0.876,-1.268"]
ave_npp [pos="-0.927,0.696"]
ave_temp [pos="-0.890,1.097"]
bed_shear_stress [latent,pos="-0.963,1.079"]
blacktip_reef_shark [pos="-0.970,-1.278"]
browser [latent,pos="-0.771,-0.705"]
cloud_cover [latent,pos="-0.785,1.464"]
common_blacktip_shark [latent,pos="-0.962,-2.393"]
coral_spawning [latent,pos="-1.250,0.392"]
crown_of_thorns [latent,pos="-1.147,-0.199"]
crustose_coraline_algae [pos="-0.923,0.349"]
cyclones [latent,pos="-1.113,1.307"]
depth [latent,pos="-0.838,1.489"]
emerged_land_area [pos="-0.883,1.585"]
grazer [pos="-0.893,-0.705"]
great_hammerhead_shark [latent,pos="-0.785,-2.397"]
grey_reef_shark [exposure,pos="-1.206,-1.756"]
hard_coral [outcome,pos="-1.030,-0.171"]
invert [pos="-1.103,0.007"]
invertivore [pos="-1.108,-0.673"]
isl_grp [pos="-1.042,1.753"]
island_geomorphology [pos="-0.970,1.724"]
lagoon_size [pos="-1.086,1.471"]
latitude [pos="-1.104,1.716"]
light [latent,pos="-0.765,1.037"]
longitude [pos="-1.090,1.884"]
nutrient_run_off [latent,pos="-1.076,1.247"]
offshore_prey [latent,pos="-1.301,-0.762"]
other_algae [pos="-0.825,0.260"]
other_offshore_prey_proxies [latent,pos="-1.283,-0.470"]
planktivore [pos="-1.203,-0.701"]
pop_dens [pos="-0.711,0.242"]
population_size [pos="-0.720,0.716"]
pursuit_piscivore [pos="-0.779,-1.271"]
relief [pos="-1.123,0.674"]
scalloped_hammerhead_shark [latent,pos="-1.092,-2.397"]
scraper [pos="-1.003,-0.698"]
season [pos="-0.720,1.543"]
sicklefin_lemon_shark [pos="-1.095,-2.097"]
silvertip_shark [latent,pos="-0.753,-1.802"]
tawny_nurse_shark [pos="-1.067,-1.285"]
tiger_shark [latent,pos="-0.879,-2.400"]
transient_pelagic_sharks [pos="-0.933,-2.078"]
turbidity [latent,pos="-0.773,0.592"]
wave_exposure [latent,pos="-0.953,1.322"]
whitetip_reef_shark [pos="-1.168,-1.385"]
zooplankton [pos="-1.210,0.096"]
ambush_piscivore -> browser
ambush_piscivore -> grazer
ambush_piscivore -> invertivore
ambush_piscivore -> planktivore
ambush_piscivore -> scraper
ave_npp -> crustose_coraline_algae
ave_npp -> hard_coral
ave_npp -> invert
ave_npp -> offshore_prey
ave_npp -> turbidity
ave_npp -> zooplankton
ave_temp -> ave_npp
ave_temp -> crustose_coraline_algae
ave_temp -> hard_coral
ave_temp -> offshore_prey
ave_temp -> other_algae
bed_shear_stress -> crustose_coraline_algae
bed_shear_stress -> relief
blacktip_reef_shark -> browser
blacktip_reef_shark -> grazer
blacktip_reef_shark -> invertivore
blacktip_reef_shark -> planktivore
blacktip_reef_shark -> scraper
browser -> other_algae
cloud_cover -> light
common_blacktip_shark -> transient_pelagic_sharks
crown_of_thorns -> hard_coral
crustose_coraline_algae -> hard_coral
cyclones -> relief
depth -> bed_shear_stress
depth -> light
depth -> wave_exposure
emerged_land_area -> nutrient_run_off
emerged_land_area -> pop_dens
grazer -> crustose_coraline_algae
grazer -> other_algae
great_hammerhead_shark -> transient_pelagic_sharks
grey_reef_shark -> ambush_piscivore
grey_reef_shark -> browser
grey_reef_shark -> grazer
grey_reef_shark -> invertivore
grey_reef_shark -> offshore_prey
grey_reef_shark -> planktivore
grey_reef_shark -> pursuit_piscivore
grey_reef_shark -> scraper
hard_coral -> coral_spawning
invertivore -> invert
invertivore -> other_algae
isl_grp -> island_geomorphology
isl_grp -> light
island_geomorphology -> ave_npp
island_geomorphology -> bed_shear_stress
island_geomorphology -> emerged_land_area
island_geomorphology -> lagoon_size
island_geomorphology -> nutrient_run_off
island_geomorphology -> offshore_prey
lagoon_size -> nutrient_run_off
latitude -> cyclones
latitude -> isl_grp
latitude -> light
light -> ave_npp
light -> ave_temp
light -> turbidity
longitude -> isl_grp
nutrient_run_off -> ave_npp
nutrient_run_off -> crown_of_thorns
nutrient_run_off -> crustose_coraline_algae
nutrient_run_off -> other_algae
nutrient_run_off -> turbidity
other_algae -> hard_coral
other_offshore_prey_proxies -> offshore_prey
planktivore -> coral_spawning
pop_dens -> ambush_piscivore
pop_dens -> browser
pop_dens -> grazer
pop_dens -> hard_coral
pop_dens -> invertivore
pop_dens -> nutrient_run_off
pop_dens -> offshore_prey
pop_dens -> planktivore
pop_dens -> pursuit_piscivore
pop_dens -> scraper
population_size -> pop_dens
pursuit_piscivore -> browser
pursuit_piscivore -> grazer
pursuit_piscivore -> invertivore
pursuit_piscivore -> planktivore
pursuit_piscivore -> scraper
relief -> browser
relief -> grazer
relief -> hard_coral
relief -> invert
relief -> invertivore
relief -> planktivore
relief -> scraper
scalloped_hammerhead_shark -> transient_pelagic_sharks
scraper -> crustose_coraline_algae
scraper -> hard_coral
scraper -> other_algae
season -> ave_temp
season -> cloud_cover
season -> light
sicklefin_lemon_shark -> ambush_piscivore
sicklefin_lemon_shark -> blacktip_reef_shark
sicklefin_lemon_shark -> grey_reef_shark
sicklefin_lemon_shark -> offshore_prey
sicklefin_lemon_shark -> pursuit_piscivore
sicklefin_lemon_shark -> silvertip_shark
sicklefin_lemon_shark -> tawny_nurse_shark
sicklefin_lemon_shark -> whitetip_reef_shark
silvertip_shark -> ambush_piscivore
silvertip_shark -> browser
silvertip_shark -> grazer
silvertip_shark -> invertivore
silvertip_shark -> offshore_prey
silvertip_shark -> planktivore
silvertip_shark -> pursuit_piscivore
silvertip_shark -> scraper
tawny_nurse_shark -> browser
tawny_nurse_shark -> grazer
tawny_nurse_shark -> invertivore
tawny_nurse_shark -> planktivore
tawny_nurse_shark -> scraper
tiger_shark -> transient_pelagic_sharks
transient_pelagic_sharks -> ambush_piscivore
transient_pelagic_sharks -> blacktip_reef_shark
transient_pelagic_sharks -> grey_reef_shark
transient_pelagic_sharks -> offshore_prey
transient_pelagic_sharks -> pursuit_piscivore
transient_pelagic_sharks -> silvertip_shark
transient_pelagic_sharks -> tawny_nurse_shark
transient_pelagic_sharks -> whitetip_reef_shark
turbidity -> crustose_coraline_algae
turbidity -> hard_coral
turbidity -> other_algae
wave_exposure -> bed_shear_stress
whitetip_reef_shark -> browser
whitetip_reef_shark -> grazer
whitetip_reef_shark -> invertivore
whitetip_reef_shark -> planktivore
whitetip_reef_shark -> scraper
zooplankton -> planktivore
}')

plot(myDag)
impliedConditionalIndependencies(myDag)
# Things involving island geomorphology:
Isl_ _||_ amb_ | pp_d
ciTest( "island_geomorphology", "ambush_piscivore", "pop_dens",
        data = ReefWideBRUVUVC,
        type = "cis.chisq") # FAILS
Isl_ _||_ amb_ | em__
ciTest("island_geomorphology", "ambush_piscivore", "emerged_land_area",
        data = ReefWideBRUVUVC,
        type = "cis.chisq") # FAILS
Isl_ _||_ av_t | isl_
Isl_ _||_ bl__
Isl_ _||_ brws | pp_d, relf
Isl_ _||_ brws | em__, relf
Isl_ _||_ grzr | pp_d, relf
Isl_ _||_ grzr | em__, relf
Isl_ _||_ gr__
Isl_ _||_ invert | av_n, invrtv, relf
Isl_ _||_ invr | av_n, pp_d, relf
Isl_ _||_ invr | pp_d, relf
Isl_ _||_ invr | em__, relf
Isl_ _||_ lttd | isl_
Isl_ _||_ lngt | isl_
Isl_ _||_ plnk | pp_d, relf, zpln
Isl_ _||_ plnk | av_n, pp_d, relf
Isl_ _||_ pp_d | em__
Isl_ _||_ ppl_
Isl_ _||_ prs_ | pp_d
Isl_ _||_ prs_ | em__
Isl_ _||_ scrp | pp_d, relf
Isl_ _||_ scrp | em__, relf
Isl_ _||_ sesn
ciTest("island_geomorphology", "season",
        data = ReefWideBRUVUVC,
        type = "cis.chisq") # FAILS
Isl_ _||_ sc__
Isl_ _||_ tw__
Isl_ _||_ tr__
Isl_ _||_ wh__
Isl_ _||_ zpln | av_n

wh__ _||_ zpln
ciTest( "whitetip_reef_shark", "zooplankton",
        data = ReefWideBRUVUVC,
        type = "cis.chisq") # FAILS
# zooplankton marked as observed in DAG but not present in dbase why?

sesn _||_ wh__
ciTest("season", "whitetip_reef_shark",
        data = ReefWideBRUVUVC,
        type = "cis.chisq") # FAILS

ciTest("site_name", "reef_name",
        data = ReefWideBRUVUVC,
        type = "cis.chisq") # FAILS

# brca.txt (download at: https://github.com/ankurankan/2020-dagitty-manual/)
brca <- read.table("/home/simon/Documents/Si Work/PostDoc Work/FIU/2024-01_SharksFishCoral-FrenchPoly/Nat resources/Ankan/brca.txt",
                   header = TRUE)
ciTest("Age", "Recurrence", c("Irradiation", "Menopause"),
       data = brca,
       type = "cis.chisq")

ciTest("geo", "island_geomorphology",
       data = ReefWideBRUVUVC,
       type = "cis.chisq") # FAILS

tmp <- ReefWideBRUVUVC |> 
  dplyr::select(geo, island_geomorphology)

ciTest("geo", "island_geomorphology",
       data = tmp,
       type = "cis.chisq") # FAILS

xtfrm.data.frame(brca$Recurrence) # fails
```

```{r load data}
ReefWideBRUVUVC <-  readr::read_csv(file =  "/home/simon/Documents/Si Work/PostDoc Work/FIU/2024-01_SharksFishCoral-FrenchPoly/NFF Data code/ReefWideBRUVUVC.csv")
# download specified DAG from dagitty.net
# DAG <- downloadGraph("dagitty.net/mt8WHQajL") # https://dagitty.net/mt8WHQajL
```

Between BRUV+UVC combined and running DAG, surely I need to get all other variables, do I have them all? See https://docs.google.com/spreadsheets/d/1x7GLAVqXHQXAk9JVcsNAaS7aMi1AQ-DO/edit#gid=750764017

See `?dagitty` for info on how to format this properly. `x [exposure,pos="1,0"]`

# define variables for protected area, fire, poaching, logging, distance to roads and cities, slope, elevation, carbon sequestration and forest species abundance

def \<- defData(varname = "slope", dist = "normal", formula = 0, variance = 1) def \<- defData(def, varname = "elevation", dist = "normal", formula = 0, variance = 1) def \<- defData(def, varname = "distancetoroadsandcities", dist = "normal", formula = 0, variance = 1) def \<- defData(def, varname = "protectedarea", dist = "binary", formula = "0.9 \* slope + 0.9 \* elevation + -1.2 \* distancetoroadsandcities", link = "logit", variance = 1) def \<- defData(def, varname = "fire", dist = "normal", formula = "-0.5 \* protectedarea + 0.6 \* distancetoroadsandcities", variance = 1) def \<- defData(def, varname = "poaching", dist = "normal", formula = "-1.1 \* protectedarea + 0.8 \* distancetoroadsandcities", variance = 1) def \<- defData(def, varname = "logging", dist = "normal", formula = "-0.7 \* slope + -0.7 \* elevation + -1.6 \* protectedarea + 0.6 \* distancetoroadsandcities", variance = 1) def \<- defData(def, varname = "forestspeciesabundance", dist = "normal", formula = "-0.4 \* fire + -0.6 \* poaching + -0.7 \* logging", variance = 1) def \<- defData(def, varname = "carbonsequestration", dist = "normal", formula = "-0.8 \* logging + 0.4 \* forestspeciesabundance", variance = 1)

# create 10000 observations

ForestSpeciesAbundanceData \<- genData(10000, def)

# Test DAG-data consistency

# evaluate the d-separation implications of our DAG with our simulated dataset

```{r DAG data consistency}
test <- localTests(x = myDag,
                   data = ReefWideBRUVUVC,
                   type = "cis",
                   abbreviate.names = FALSE)
# Warning in cov2cor(cov(data)): diag(.) had 0 or NA entries; non-finite result is doubtful
# Error in sample.cov[vars, vars] : subscript out of bounds
# > traceback()
# 7: .ci.test.covmat(sample.cov, sample.nobs, i, conf.level, tol)
# 6: FUN(X[[i]], ...)
# 5: lapply(X = X, FUN = FUN, ...)
# 4: sapply(tests, f)
# 3: t(sapply(tests, f))
# 2: as.data.frame(row.names = row.names, t(sapply(tests, f)))
# 1: localTests(x = myDag, data = ReefWideBRUVUVC, type = "cis")

print(test)
```

# perform Holm-Bonferrino correction to mitigate problems around multiple testing

test$p.value <- p.adjust(test$p.value)

test \# should show all p values above 0.05, suggesting DAG-data consistency
